---
title: "Lab 6"
author: "Hunter Finger"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:**  4 (Pg. 302); 1 (Pgs. 316-317); 1 (Pgs. 328-329); 1, 2 (Pgs. 353-354)

**Assigned:** Friday, March 22, 2019

**Due:** Friday, March 29, 2019 by 5:00 PM

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining.

```{r, include=FALSE}
library(tidyverse)
```

# Chapter 16 (Pg. 302)

##  Exercise 4

### a)
```{r}
last_value = function(x){
    x[[length(x)]]
}
last_value(1:8)
last_value(1)
  
```

This function uses the double brackets to extract a single value.

### b)
```{r}
even_pos = function(x){
    x[seq_along(x) %% 2 == 0]
}

even_pos(1:13)
```

### c)
```{r}
except_last = function(x){
  x[1:length(x)-1]
}

except_last(1:5)
except_last(1)
```

### d)
```{r}
even_num = function(x){
  x[x %% 2 == 0]
}

even_num(8:13)
```

# Chapter 17 (Pgs. 316-317)

##  Exercise 1

### a)
```{r}
answer <- vector("double", ncol(mtcars))
names(answer) = names(mtcars)
for (i in names(mtcars)){
  answer[i] = mean(mtcars[[i]])
}
answer
```

### b)
```{r}
answer = vector("list", ncol(nycflights13::flights))
names(answer) = names(nycflights13::flights)
for (i in names(nycflights13::flights)){
  answer[i] = class(nycflights13::flights[[i]])
}
answer
```

### c)
```{r}
answer = vector("double", ncol(iris))
names(answer) = names(iris)
for(i in names(iris)){
  answer[i] = length(unique(iris[[i]]))
}
answer
```

### d)
```{r}
n <- 10
av <- c(-10, 0, 10, 100)
normals <- vector("list", length(av))
for (i in seq_along(normals)) {
  normals[[i]] <- rnorm(n, mean = av[i])
}
normals
```

# Chapter 17 (Pgs. 328-329)

##  Exercise 1

### a)
```{r}
map_dbl(mtcars, mean)
```

### b)
```{r}
map_chr(nycflights13::flights, typeof)
```

### c)
```{r}
map_int(iris, function(x) length(unique(x)))
```

### d)
```{r}
map(c(-10, 0, 10, 100), ~ rnorm(n = 10, mean = .))
```

# Chapter 18 (Pgs. 353-354)

##  Exercise 1
```{r}
sim1a <- tibble(x = rep(1:10, each = 3), y=x*1.5+6+rt(length(x),df=2))
ggplot(sim1a, aes(x=x,y=y)) + geom_point() + geom_smooth(method = "lm", se = F)
```

##  Exercise 2
```{r}
measure_distance <- function(mod, data) {
  diff <- data$y - make_prediction(mod, data)
  mean(abs(diff))
}

make_prediction <- function(mod, data){
  mod[1] + mod[2] * data$x
}

model <- optim(c(0,0), measure_distance, data = sim1a)
model
```

The slope and y intercept are very close to what you see in exercise 1 of this chapter. Optim is essentially doing the same thing that the linear model did with similar results.