---
title: "Lab 1"
author: "Hunter Finger"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:** 1-5 (Pgs. 6-7); 1-2, 5 (Pg. 12); 1-5 (Pgs. 20-21); Open Response

**Assigned:** Friday, January 11, 2019

**Due:** Friday, January 18, 2019 by 5:00 PM

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a html file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. 

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
```

# Chapter 1 (Pgs. 6-7)

##  Exercise 1
```{r}
ggplot(data=mpg)
```

I see a box where a graph could be.

## Exercise 2
```{r}
dim(mtcars)
nrow(mtcars)
ncol(mtcars)
```
There are 32 rows and 11 columns in the dataset **mtcars**.

##  Exercise 3
```{r}
?mpg
```

```{r}
unique(mpg$drv)
```


The variable **drv** is a factor variable that takes the following values:

- "f" = front-wheel drive
- "r" = rear-wheel drive
- "4" = 4-wheel drive

##  Excercise 4
```{r}
ggplot(data=mpg,aes(x=hwy,y=cyl)) +
  geom_point() + 
  xlab("Highway Miles Per Gallon") +
  ylab("Number of Cylinders")
```


##  Excercise 5
```{r}
ggplot(data=mpg,aes(x=class,y=drv)) + 
  geom_point() + 
  xlab("Type of Car") +
  ylab("Type of Drive")
```

Scatter plots are not meant to visualize the relationship between two categorical/qualitative variables. 

# Chapter 1 (Pg. 12)

##  Exercise 1
The points are not blue because the color declaration needs to be outside of the aes tag.
```{r}
ggplot(data=mpg)+geom_point(mapping=aes(x=displ, y=hwy), color = "blue")
```

##  Exercise 2
```{r}
?mpg
```
**manufacturer**, **model**, **trans**, **drv**, **fl**, and **class** are all categorical. The only continuous variable is **displ**.

##  Exercise 5
```{r}
ggplot(data=mpg)+geom_point(mapping = aes(x=displ, y=hwy, stroke=cyl))
```

The **stroke** controls how large the points are based on how many cylinders (**cyl**) the engine has. The smaller the cylinder the smaller the size of the dot.

# Chapter 1 (Pgs. 20-21)

##  Exercise 1
To draw a boxplot, **geom_boxplot** would be used. For a histogram, **geom_histogram** would be used. For a line graph, **geom_line** would be used. For an area chart, **geom_area** is used.
```{r}

```

##  Exercise 2

I think that this code is going to show the car's engine liter vs highway mpg. The different colors are going to show the drivetrains of the cars and a regression line will be added by drivetrain. The **se=F** will do something, but I do not know what. 
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv))+ geom_point() + geom_smooth(se=F)
```

My prediction was correct and the **se=F** takes the confidence interval off of the regression lines.

##  Exercise 3

**show.legend = FALSE** make the legend on the side of the graphs go away. It was used earlier in the chapter to be able to have the graphs side by side without the legend being repetitive and taking up space. 

```{r}
#
```

##  Exercise 4

As I discovered in the second exercise, **se** is the confidence interval that is around the lines (as shown in the bottom example)

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv))+ geom_point() + geom_smooth(se=F)
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv))+ geom_point() + geom_smooth()
```

##  Exercise 5

I don't know if they will look different. Let me check.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

They do not look different. I am incredibly surprised.

# Open Response

For this exercise, use the **diamonds** dataset in the tidyverse. Use `?diamonds` to get more information about the dataset.

## Step 1: Select 1 numeric variable and 2 categorical variables. Create a graphic using `geom_boxplot()` and `facet_wrap` to illustrate the empirical distributions of the sample.

```{r}
ggplot(data = diamonds)+
  geom_boxplot(mapping = aes(x = clarity, y = price, color = clarity))+
  facet_wrap(~ color, nrow = 3)+
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) + 
  guides(color=guide_legend(title="Clarity"))+
  xlab("") + ylab("Price (Dollars)")
```

## Step 2: Choose 2 numeric variables and 2 categorical variables and creatively illustrate the relationship between all the variables. 

```{r}
ggplot(data = diamonds)+ geom_point(mapping = aes(x = carat, y = price, color = clarity, shape = cut))
```